---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# microflowseq

<!-- badges: start -->
[![R-CMD-check](https://github.com/PennChopMicrobiomeProgram/microflowseq/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PennChopMicrobiomeProgram/microflowseq/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/PennChopMicrobiomeProgram/microflowseq/branch/main/graph/badge.svg)](https://app.codecov.io/gh/PennChopMicrobiomeProgram/microflowseq?branch=main)
<!-- badges: end -->

The goal of microflowseq is to assign probabilities to bacteria that are sorted
in mFLOW-Seq experiments.

## Installation

You can install the development version of microflowseq like so:

``` r
devtools::install_github("PennChopMicrobiomeProgram/microflowseq")
```

## Example: dual-seq

```{r include=FALSE}
devtools::load_all()
```

```{r eval=FALSE}
library(microflowseq)
```

```{r}
library(tidyverse)
```

```{r}
# Species in rows, channel in columns
dual_cts <- cbind(`IgA+` = c(sa = 5, sb = 7), `IgA-` = c(sa = 19, sb = 13))
```

The real answer

```{r}
dual_cts %>%
  sweep(1, rowSums(dual_cts), `/`)
```

The things we measure

```{r}
dual_fracs <- renormalize(colSums(dual_cts))
# Species in rows, channels in columns
dual_props <- dual_cts %>%
  sweep(2, colSums(dual_cts), `/`)
```

Solve in matrix format

```{r}
mflow_apply(dual_props, dual_fracs)
```

Solve in long format. First prepare the data.

```{r}
dual_props_df <- dual_props %>%
  as.data.frame() %>%
  rownames_to_column("asv") %>%
  pivot_longer(-asv, names_to = "fraction", values_to = "proportion") %>%
  # In real studies, we'll have more specimens
  # Add a specimen ID so we can demonstrate how it would work
  mutate(specimen_id = "Specimen1") %>%
  select(specimen_id, everything())
dual_fracs_df <- dual_fracs %>%
  enframe("fraction", "fraction_abundance") %>%
  mutate(specimen_id = "Specimen1") %>%
  select(specimen_id, everything())
```

Then, use the tidyverse to generate prbabilities.

```{r}
dual_props_df %>%
  # Join the table of fraction abundances for each specimen
  left_join(dual_fracs_df, by = c("specimen_id", "fraction")) %>%
  # The function is valid for one ASV in one specimen
  group_by(specimen_id, asv) %>%
  # I'm tempted to name the column the same as the function :(
  mutate(mflow_prob = mflow_probability(proportion, fraction_abundance))
```

## Example: multi-seq

```{r}
# Species in rows, channel in columns
multi_cts <- cbind(
  `IgA+IgG+` = c(sa = 5, sb = 7),
  `IgA+IgG-` = c(sa = 19, sb = 13),
  `IgA-IgG+` = c(sa = 2, sb = 11),
  `IgA-IgG-` = c(sa = 23, sb = 29))
```

The real answer

```{r}
multi_cts %>%
  sweep(1, rowSums(multi_cts), `/`)
```

The things we measure

```{r}
multi_fracs <- renormalize(colSums(multi_cts))
# Species in rows, channels in columns
multi_props <- multi_cts %>%
  apply(2, renormalize)
```

Solve in matrix format

```{r}
mflow_apply(multi_props, multi_fracs)
```
