---
title: "deconvolution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{deconvolution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
devtools::load_all()
```

```{r eval=FALSE}
library(microflowseq)
```

Following the deconvolution example at
https://www.nature.com/articles/s41591-019-0480-9/figures/7

The correct answer for the count of balls that are pink and blue, large and
small.

```{r}
pb_correct <- cbind(
  small = c(pink = 12, blue = 11),
  large = c(pink = 6, blue = 5))
pb_correct
```

The large and small balls aren't sorted perfectly. Here are the pink and blue
balls that we measure in each fraction.

```{r}
pb_sorted <- cbind(
  small.fraction = c(pink = 11, blue = 10),
  large.fraction = c(pink = 7, blue = 6))
pb_sorted
```

We can sort again to measure how many small and large balls were in each
fraction.

```{r}
pb_postsep <- rbind(
  small.fraction = c(small = 18 / 21, large = 3 / 21),
  large.fraction = c(small =  5 / 13, large = 8 / 13))
pb_postsep
```

In the figure, they show us how to solve for the absolute number of pink and
blue balls in each fraction. Let's compute that and compare to the correct
answer.

```{r}
pb_sorted %*% pb_postsep
pb_correct
```
In an experiment, we measure the relative abundance of pink and blue balls in
each fraction. Here are the relative abundances we measure in each fraction.

```{r}
pb_sorted_relabund <- apply(pb_sorted, 2, normalize_sum)
pb_sorted_relabund
```

Solve for the relative abundance of balls and compare to the correct answer.
Success!

```{r}
apply(pb_sorted_relabund %*% pb_postsep, 2, normalize_sum)
apply(pb_correct, 2, normalize_sum)
```

